---
title: "BellabeatAnalysis_GoogleCapstone"
author: "Data Nathan V"
date: "`r Sys.Date()`"
output: html_document
---

# Introduction and Background

Bellabeat is a tech company that produces health-focused smart products focused on women. Founded by Urška Sršen and Sando Mur in 2013, the company collects data from their technology to empower women with knowledge about their own health and habits. Bellabeat's website can be found [here.](https://bellabeat.com/)

This project is for the completion of my Google Data Analytics Capstone. 


# Business Task
Identify trends and possible opportunities for Bellabeat using data from a different smart device, FitBit Fitness Tracker Data.


# Hypotheses
These are my hypotheses from the data:

* Total steps is directly proportional to calories burnt.
* Time spent sedentary is inversely proportional to calories burnt.
* Time spent in bed is directly proportional to total minutes of sleep.
* Total minutes of rest is directly proportional to calories burnt.


# Data Source
This dataset includes thirty consenting FitBit users to use their personal data on minute-level output for physical activity, heart rate, sleep monitoring. The data is available on Kaggle, a public domain and open-source repository for users to share data. Click [here](https://www.kaggle.com/datasets/arashnic/fitbit) for the complete dataset and more information.


# Documentation
Firstly, I have to load the packages I need for this project.
```{r}
library(tidyverse)
library(tidyr)
library(dplyr)
library(skimr)
library(janitor)
library(lubridate)
library(ggplot2)
```

Next step would be importing the data I need for this analysis
```{r}
dailyactivity <- read.csv("Fitabase Data 4.12.16-5.12.16/dailyActivity_merged.csv")
sleep <- read.csv("Fitabase Data 4.12.16-5.12.16/sleepDay_merged.csv")
```

Now, I want to check the summaries of the data that I will be using.
```{r}
head(dailyactivity)
colnames(dailyactivity)
str(dailyactivity)
head(sleep)
colnames(sleep)
str(sleep)
```

From these summaries, the **ID** column would be our common denominator to merge both tables. Now I want to know how many unique participants in both **dailyactivity** and **sleep** dataframes.
```{r}
n_distinct(dailyactivity$Id)
n_distinct(sleep$Id)
```

Given that, it shows that there are 33 participants in **dailyactivity** and 24 in **sleep**. 

I want to merge both dataframes so I can start my analysis. I will name it **activity_sleep_merged**
```{r}
activity_sleep_merged <- merge(dailyactivity, sleep, by=c("Id"))
```
Then I want to verify the new total number of participants for my clean data.
```{r}
n_distinct(activity_sleep_merged$Id)
```
Next would be to check the summary of my merged data.
```{r}
head(activity_sleep_merged)
colnames(activity_sleep_merged)
str(activity_sleep_merged)
```
Lastly, I want to group the data by **Id** using the average for **total steps**, **calories**, **time spent  sedentary**, **time in bed**, and **total time asleep**.
```{r}
groupave_steps_calories_sedentary_bed_sleep <- 
  activity_sleep_merged %>% 
  group_by(Id) %>% 
  summarise(Steps = mean(TotalSteps), Cal = mean(Calories), Sedentary = mean(SedentaryMinutes), Bedtime = mean(TotalTimeInBed), Sleep =   mean(TotalMinutesAsleep))
```
I would want to check the summary of this dataframe
```{r}
head(groupave_steps_calories_sedentary_bed_sleep)
colnames(groupave_steps_calories_sedentary_bed_sleep)
str(groupave_steps_calories_sedentary_bed_sleep)
```


Now I can start my analysis.

# Analysis
I want to start with some summary statistics from the clean data. I will get the average together with standard deviation for total steps, calories, sleep, and time in bed.

```{r}
groupave_steps_calories_sedentary_bed_sleep %>% 
  summarise(mean(Steps), sd(Steps), mean(Cal), sd(Cal), mean(Sedentary), sd(Sedentary), mean(Bedtime), sd(Bedtime), mean(Sleep), sd(Sleep))
```

## Total steps is directly proportional to calories burnt.
I want to see if there is a positive relationship between total steps and calories burnt. I will start by using the correlation function. I am making the assumption that total steps in general is correlated to higher calories burnt.

```{r}
groupave_steps_calories_sedentary_bed_sleep %>% summarise(cor(Steps, Cal))
```

Now I want to plot the relationship between total steps and calories.

```{r}
ggplot(data=groupave_steps_calories_sedentary_bed_sleep, mapping = aes(x=Steps, y=Cal)) + geom_point() + geom_smooth(color="green") + labs(title = "Steps vs. Calories Burnt")
```

Based from this analysis, there is a 0.322 correlation or 32.2%.

## Time spent sedentary is inversely proportional to calories burnt.
I want to see if there is a negative relationship between time sedentary and calories burnt. I will start by using the correlation function.

```{r}
groupave_steps_calories_sedentary_bed_sleep %>% summarise(cor(Sedentary, Cal))
```

Now I want to plot the relationship between time sedentary and calories.

```{r}
ggplot(data=groupave_steps_calories_sedentary_bed_sleep, mapping = aes(x=Sedentary, y=Cal)) + geom_point() + geom_smooth(color="red") + labs(title = "Sedentary Minutes vs. Calories Burnt")
```

Based from this analysis, there is a -0.0545 correlation or -5.45%.

## Time spent in bed is directly proportional to total minutes of sleep.
I want to see if there is a positive relationship between time in bed and total minutes of sleep. I will start by using the correlation function.

```{r}
groupave_steps_calories_sedentary_bed_sleep %>% summarise(cor(Bedtime, Sleep))
```

Now I want to plot the relationship between time in bed and total time asleep.

```{r}
ggplot(data=groupave_steps_calories_sedentary_bed_sleep, mapping = aes(x=Bedtime, y=Sleep)) + geom_point() + geom_smooth(color="green") + labs(title = "Time in Bed vs. Total Time Asleep")
```

Based from this analysis, there is a 0.940 correlation or 94%.

## Total minutes of rest is directly proportional to calories burnt.
I want to see if there is a positive relationship between total minutes of sleep and calories. I will start by using the correlation function.

```{r}
groupave_steps_calories_sedentary_bed_sleep %>% summarise(cor(Sleep, Cal))
```

Now I want to plot the relationship between time sedentary and calories.

```{r}
ggplot(data=groupave_steps_calories_sedentary_bed_sleep, mapping = aes(x=Sleep, y=Cal)) + geom_point() + geom_smooth(color="red") + labs(title = "Total Time Asleep vs. Calories Burnt")
```

Based from this analysis, there is -0.0406 correlation or -4.06%.


# Summary and Recommendations

The last part of this case study would be focused on key findings and how the Bellabeat App can capitalize on several opportunities

I want to start with the most significant result of the research. Total time in bed is highly proportional to total time in bed. I would suggest that the Bellabeat App to enable users to have the option of having notifications when it detects sleep patterns like common bedtime. The app can show how many hours of sleep that the user can get for that notification. The app can also feature a "Do not disturb" mode so they can also get a better quality of sleep. 

The second recommendation would be on total steps and total calories burnt. With the 32% correlation based on our findings, it is an opportunity to bump this rate by reminding the users of their daily step count. Bellabeat can also send out notifications on articles and studies on how having an active lifestyle benefits their users. The app can also feature a social aspect where users can compare their step count with their friends or an option to share their steps and if they achieved a new goal. 

My next point would be the relationship between time spent sedentary and calories burnt. I made the assumption that less calories will be used whenever the user spent more time being sedentary. My findings show a -5.45% correlation, which was expected but I thought the margin would be greater than that. It should also be noted that I did not take into account other metrics like how active they are for the calories spent. This is still an opportunity for Bellabeat regarding the second recommendation.

Lastly, the relationship between total time asleep and calories burnt should be discussed. My assumption was that users would be able to burn more calories if they get more rest. The result shows a -4.06% which invalidates my assumption. It is a very small margin just like the last point. 


The clean sample size of 24 is not sufficient in my opinion but there are significant findings from the analyses. Furthermore, I used the average for all the categories for each users, within the two months that data was collected. I highly suggest having a larger sample size and a longer time period to gather data for further researcher. 



